### 2019-09-24 response_to_referees.txt
added old response to referee

Many thanks to the referee for their helpful and considerate suggestions.

"the relationship to weighted (co)limits should be clarified"
-> We have added Remark 4.4 on the referee's question about weighted colimits.

"It seems that the correct spelling is "Karvonen""
-> We have corrected 'Korvanen' to 'Karvonen'.

"I see the point that analytical structures aren't usually understood in terms of universal properties, but the sentence in quotes sounds a bit bizarre - how about geometric structures?"
-> We have rephrased the sentence.

"It would be good for the authors to give the definitions of a projection, isometry, unitary."
-> We have added definitions of isometry and unitary to Definition 2.2. Our proof of Lemma 3.2 was overkill, in that b*b = 1 is enough, and we have removed the extra unnecessary part.

"would read better as "there is a _morphism_ b:A->A...""
-> Fixed.

"Proof of Lemma 3.2: I don't understand how the reasoning..."
-> The reviewer tried to pattern match the wrong thing. We added "... = 1-1 = 0 ≤ 0".

"The word order in "For A_i a  family of objects" would be improved by e.g. "For a family A_i of objects in C.."" or "For a family of objects A_i in C...""
-> Fixed.

"please derive explicitly the triangle equality for the norm"
->  We reformulated the text to make more clear that this is a standard result known from the theory of Hilbert C*-modules. We also added a second citation to Paschke’s original proposition.

"might read better by adding an "i.e.", as in  "is uniformly bounded by some constant C, i.e.""
-> Fixed.

"Typo: should be "The covergence f^{(n)}->f works similarly" rather than "f_i->f""
-> Fixed.

"it might be better to say "for suffienctly large n""
-> We have followed the referee's suggestion of using "for sufficiently large n".

"However, it would make the readers life easier to write down the easy proof"
-> We have added some explanation, now on top of p.7.

"no need to try, just do"
-> Fixed, we just do.

"them 5.1(b) is poorly worded"
-> Fixed.

"proof of (b)->(c)"
-> Yes, corrected.

"Maybe this is standard, but it hurts my eyes"
-> Fixed.

"ref [4] has been published in TAC, please cite it appropriately."
-> Fixed.